An Efficient Forecasting Approach\\ to Reduce Boundary Effects in Real-Time Time-Frequency Analysis

### 2020-09-03 Notes/BoundEffRed.tex
update paper

Various improvements can be considered to make the algorithm more robust. In particular, we have noticed (see Fig.~\ref{fig:THO.failure}) that when the regular oscillations of the observed signal break, the forecasting step is no longer relevant, and only slows down the calculation of the TF representation. A preliminary step should then be added to the algorithm to detect signal activity, and disable the forecasting step when possible. A more fundamental improvement would be to perform the forecasting step directly in the TF plane, to specifically predict the evolution of each instantaneous frequency. This leads to seeing a TF representation as a multivariate time series, with each frequency band containing a univariate time series. These avenues will be explored in our future work.

>>>>

Various improvements can be considered to make the algorithm more robust. In particular, we have noticed (see Fig.~\ref{fig:THO.failure}) that when the regular oscillations of the observed signal break, the forecasting step is no longer relevant, and only slows down the calculation of the TF representation. A preliminary step should then be added to the algorithm to detect signal activity, and disable the forecasting step when possible. More fundamentally, one can also consider accelerating the computational time by optimizing the forecasting step, \ie~reduce $t_{\mathrm{forecast}}$. Indeed, each new forecast requires the inversion of the matrix $\bX\bX^T$ of size $M\times M$. However, the matrix $\bX$ used to forecast $\bx_{N+H}$ at the current iteration differs from the one used at the previous iteration to forecast $\bx_{N}$ by only $H$ columns. It then seems natural to take inspiration from the work of Strobach~\cite{Strobach97square}, generalized by Badeau \etal~\cite{Badeau04sliding}, which proposes a fast algorithm for the singular value decomposition of successive data matrices taking the same form as $\bX$. Their algorithm is limited to the case where $H=1$. Developing an extension of this algorithm to variations of $H>1$ columns would then allow to efficiently update the matrix $\left(\bX\bX^T\right)^{-1}$. These avenues will be explored in our future work.

### 2020-09-03 Notes/results.tex
update paper

First, when the noise is weak, the variance of the forecasting estimator is dramatically high. For instance, in this specific example, the estimator $\tilde\bx[N-1+\ell]$ for $\ell=10$ is inaccurate as soon as the noise variance $\sigma^2$ is lower than $4\times 10^{-14}$ (see the red line in Fig.~\ref{fig:res.noise.sine}). This validates the result of Theorem~\ref{th:error}, highlighting the necessary presence of noise to ensure the numerical stability of the estimator. Furthermore, as soon as $\sigma$ is higher than this critical threshold, the estimator variance increases linearly with $\sigma^2$, as predicted by the bound~\eqref{eq:cov.error.2} provided in Theorem~\ref{th:error}. Note that the parameter $\ell$ has little influence on the quality of the estimator. This is due to the \textit{stationarity} of the studied signal, allowing long-range forecasts without deterioration of the estimator.

@@

As anticipated by Theorem~\ref{th:error}, the variance of the estimator decreases linearly with $\frac1K$ as long as the dataset size is large enough. Indeed, in the log-log plot displayed in Fig.~\ref{fig:res.size.sine}, as soon as $K>700$, each of the three curves representing the forecasting variance tends towards an asymptote of slope $-1$.

%\paragraph{Summary}
%Both previous experimental results combined with the theoretical asymptotic equation~\eqref{eq:cov.error.2} allow us to describe the influence of the noise variance and the size of the training dataset on the variance of the forecasting noise, empirically summarized as follows:
%\begin{equation}
%\gamma[N-1+\ell,N-1+\ell] \underset{K\to\infty}{\approx} \dfrac{\sigma^2}{K}g[\ell] .
%\end{equation} 
%where $g$ is a bounded positive function. The empirical result is coherent with the theoretical result provided by Theorem~\ref{th:error}.

>>>>

First, when the noise is weak, the variance of the forecasting estimator is dramatically high. For instance, in this specific example, the estimator $\tilde\bx[N-1+\ell]$ for $\ell=10$ is inaccurate as soon as the noise variance $\sigma^2$ is lower than $4\times 10^{-14}$ (see the red line in Fig.~\ref{fig:res.noise.sine}). This validates the result of Theorem~\ref{th:error}, highlighting the necessary presence of noise to ensure the numerical stability of the estimator. Furthermore, as soon as $\sigma^2$ is higher than this critical threshold, the estimator variance increases linearly with $\sigma^2$, as predicted by the bound~\eqref{eq:cov.error.2} provided in Theorem~\ref{th:error}. Note that the parameter $\ell$ has little influence on the quality of the estimator. This is due to the \textit{stationarity} of the studied signal, allowing long-range forecasts without deterioration of the estimator.

@@

As anticipated by Theorem~\ref{th:error}, the variance of the estimator decreases linearly with $\frac1K$ as long as the dataset size is large enough. Indeed, in the log-log plot displayed in Fig.~\ref{fig:res.size.sine}, as soon as $K>700$, each of the three curves representing the forecasting variance in function of $K$ tends towards an asymptote of slope $-1$.
